---
title: "Titanic-Tutorial"
author: "Courtney Ferguson Lee"
date: "5/1/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r cars}
library(ggplot2)
library(ggthemes)
library(scales)
library(dplyr)
library(tidyr)
library(mice)
library(swirl)
library(randomForest)
library(readr)
```

## Reading Data

Note: bind_rows is a way of combining two data frames that match up all variables.  This is different from join() in that it seems the dataframes need to match up exactly.

```{r}

setwd('/Users/courtneyfergusonlee/kaggle/titanic')
train <- read.csv('train.csv', stringsAsFactors = F)
test <- read.csv('test.csv', stringsAsFactors = F)

?read.csv

str(train)
str(test)

# joins the train and test dataframes into a larger one
full <- bind_rows(train, test)

str(full)
```

## Feature Engineering

This section takes explicit variables and generates new, implicit ones from the information.  The double backslash was confusing at first.  Since \ itself needs
to be escaped in R, we need to escape these metacharacters with double backslash
like \\$ or \\.


```{r}
full$Name

head(full$Name)
grep("(.*, )", full$Name, value = T)

# This searches for and replaces two parts:
# 1. Anything before and including the comma
# 2. Anything after and including the period
full$Title <- gsub('(.*, )|(\\..*)', '', full$Name)

# This creates a table of titles and genders.
# It gives a count of every title by gender
table(full$Sex,full$Title)
```


```{r}
# Create a list of rare titles
rare_title <- c('Don', 'Dona', 'Lady', 'the Countess', 'Capt', 'Col', 
                'Dr', 'Major', 'Rev', 'Sir', 'Jonkheer')

# This looks for and replaces Mlle, Ms, Mme and rare titles
full$Title[full$Title == 'Mlle'] <- 'Miss'
full$Title[full$Title == 'Ms'] <- 'Miss'
full$Title[full$Title == 'Mme'] <- 'Mrs'
full$Title[full$Title %in% rare_title] <- 'Rare Title'

# Check the new table of Titles
table(full$Sex,full$Title)
```

Wanted to try using regular expressions on finding the surname but it doesn't 
work the way it does in python.

```{r}

# Need to use sapply here because using mutate would end up with issues... Array indexing...
full$Surname <- sapply(full$Name,  
                      function(x) strsplit(x, split = '[,.]')[[1]][1])

```


```{r}
# paste() concatenates vectors after converting to a character
# cat outputs to the console apparently

cat(paste('We have <b>', nlevels(factor(full$Surname)), '</b> unique surnames.  I would be interested to infer ethnicity based on surname --- another time.'))
?cat
```

### Add family size variable

Now that we’ve taken care of splitting passenger name into some new variables, we can take it a step further and make some new family variables. First we’re going to make a family size variable based on number of siblings/spouse(s) (maybe someone has more than one spouse?) and number of children/parents.

```{r}
# Number of family members includes siblings, spouses, parents, children and the individual
full$Fsize <- full$SibSp + full$Parch + 1

# Last name and number of family members
full$Family <- paste(full$Surname, full$Fsize, sep = '_')

```


```{r}

# Box plot of survival broken down by family size
ggplot(subset(full, !is.na(Survived)), 
       aes(x = Fsize, 
           fill = factor(Survived))) +
  geom_bar(position = 'dodge') +
  scale_x_continuous(breaks = c(1:11)) +
  labs(x = 'Family Size') +
  theme_few()

```


```{r}

# Break family size into 3 categories
full$FsizeD[full$Fsize == 1] <- 'singleton'
full$FsizeD[full$Fsize <= 4 & full$Fsize > 1] <- 'small'
full$FsizeD[full$Fsize > 4] <- 'large'

# Show family size by survival using a mosaic plot
mosaicplot(table(full$FsizeD,
                 full$Survived),
           main = 'Family Size by Survival',
           shade = T)
?mosaicplot
```

# Cabin Data

```{r}

full$Cabin[1:28]

parse_character(full$Cabin)
parse_character('C38')
full[28, ]
```


```{r}
full$Deck <- sapply(full$Cabin, function(x) strsplit(x, NULL)[[1]][1])
head(full$Deck, 20)

```


```{r}
ggplot(aes(x = Deck),
       data = subset(full, !is.na(Deck) & !is.na(Survived))) +
  geom_bar() +
  facet_wrap(~ Survived)

?facet_wrap
```

## Missingness

### Sensible Value Imputation

```{r}
full[c(62, 830), ]
```


```{r}
cat(paste('We will infer their values for **embarkment** based on present data that we can imagine may be relevant: **passenger class** and **fare**. We see that they paid<b> $', full[c(62, 830), 'Fare'][[1]][1], '</b>and<b> $', full[c(62, 830), 'Fare'][[1]][2], '</b>respectively and their classes are<b>', full[c(62, 830), 'Pclass'][[1]][1], '</b>and<b>', full[c(62, 830), 'Pclass'][[1]][2], '</b>. So from where did they embark?'))
```

This creates a boxplot based on where they embarked, what passenger class they were (1st, 2nd, 3rd), and how much their fare was.

```{r}
# Create a set that has embarkment data (get rid of 62 and 830)
full.embark_fare <- full %>%
  filter(PassengerId != 62 & PassengerId != 830)

# Use ggplot2 to visualize embarkment, passenger class & median fare
ggplot(full.embark_fare, 
       aes(x = Embarked, 
           y = Fare,
           fill = factor(Pclass))) +
  geom_boxplot() +
  geom_hline(aes(yintercept = 80),
                 color = 'red',
                 linetype = 'dashed',
                 lwd = 2) +
  scale_y_continuous(labels = dollar_format()) +
  theme_few()
```

Based on the graphs, we can see the price paid for the missing values ($80) coincides with Charbourg ('C').

```{r}
full[c(62, 830), 'Embarked'] <- 'C'

```


```{r}
full[1044, ]
```

This creates a histogram with a much smoother line.  Add this to the toolbelt.

```{r}

ggplot(subset(full, Pclass == 3 & Embarked == 'S'),
       aes(x = Fare)) +
  geom_density(fill = '#99d6ff') +
  #geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 70, 5), limits = c(0, 70)) +
  geom_vline(aes(xintercept = median(Fare, na.rm = T)),
                 color = 'red',
                 linetype = 'dashed',
                 lwd = 1)

```

Considering the graph, it seems reasonable to replace the missing fare with the median fare of $8.05

```{r}
full$Fare[1044] <- median(full[full$Pclass == 3 & full$Embarked =='S',]$Fare, na.rm = T)
```


### Predictive Imputation

There are 263 missing age values.  We're going to create a model to predict ages based on other variables.

```{r}
# Show number of missing Age values
sum(is.na(full$Age))
```

We're going to use the mice (multiple imputation using chained equations) package to predict missing ages. We're going to factorize the factor variables, then perform mice imputation. 

```{r}

# Make variable factors into factors
factor_vars <- c('PassengerId', 'Pclass', 'Sex', 'Embarked',
                 'Title', 'Surname', 'Family', 'FsizeD')

# Use lapply to turn each factor variable into a factor
full[factor_vars] <- lapply(full[factor_vars], function(x) as.factor(x))

# Set a random seed
set.seed(129)
```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```

